---
title: "Random Forest, Cycle Prediction"
output:
  html_document:
    df_print: paged
---

The goal of this project is to be able to predict the onset of a ovulation cycle, given the termination point (date) of the prior ovulation cycle, i.e. the actual length of a cycle, given the past history of he patient.  The valuation of this process is measure by how much been a prediction could be over a naive estimate.  The naive estimate, ignore all other attributes is based upon the mean cycle length, which based on the current data set is 28.75 days.   

```{r}
# Libraries 
library(readr)
library(randomForest)
# Script to predict end dates 
set.seed(1234)
```

##Step 1, Load the Data Set 

```{r}
data <- read_csv("~/GitHub/PhillyAI_Hackathon_SingleSyllable/data_out/model_sample_02.csv")
```

##Step 2, Calculate the naive Estimate 

```{r}
# Naive model 
ave_cycle = mean(data$TARGET_act_cycle)
print(ave_cycle)
```

##Step 3, Created the train/test data sets (Partition)

```{r}
# create test and training sets 
inTrain <- sample(1:nrow(data), nrow(data) * 0.85) # select 85% of the items 
train <- data[inTrain, ]
test <- data[-inTrain, ]
rm(data, inTrain)
```

##Step 4, Build Model and Make Predicitions 

```{r}
cols <- c('user_id', 'acne', 'backache',
          'bloating', 'cramp', 'diarrhea', 
          'dizzy', 'headache', 'mood', 'nausea', 
          'sore', 'est_cycle', 'est_period', 'act_period')
rf <- randomForest(train$TARGET_act_cycle ~ ., data=train[,cols], ntree=200)
predicted <- predict(rf, test[,cols])
```

##Step 5, Determine Accuracy of RF model using Root-Mean-Square-Error (RMSE)

```{r}
# actual results 
results <- as.data.frame(cbind(test$TARGET_act_cycle, predicted))
results$err = (results$V1-results$predicted)^2
mse_rf = mean(results$err)
print(sqrt(mse_rf))
```

##Step 6, Determine Accuracy of naive Model using RMSE 

```{r}
# Naive Results 
results2 <- as.data.frame(cbind(test$TARGET_act_cycle))
results2$predicted = ave_cycle
results2$err = (results2$V1-results2$predicted)^2
mse_ne = mean(results2$err)
print(sqrt(mse_ne))

```

##Step 7, Compare accuracy of RF and Naive models 

```{r}
# Improvement 
net_improvement = (mse_ne - mse_rf)/mse_ne
print(net_improvement)
```

##Conclusion 

Based upon the evidence provided, it appears that the RF model improved accuracy over a naive model by approximately 21.47%. 




